A publishing infrastructure for AI-assisted academic authoring

### 2023-12-19 latex/manuscript.tex
latex: update manuscript.tex

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A publishing infrastructure for AI-assisted academic authoring},
  pdfauthor={Milton Pividori, Casey S. Greene},
  pdflang={en-US},
  pdfkeywords={manubot, artificial intelligence, scholarly publishing, software},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\begin{itemize}
\item
  \textbf{Milton Pividori}
  \includesvg[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0000-0002-3035-4403}{0000-0002-3035-4403}
  · \includesvg[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/miltondp}{miltondp}
  · \includesvg[width=0.16667in,height=0.16667in]{images/twitter.svg}
  \href{https://twitter.com/miltondp}{miltondp}
  · \includesvg[width=0.16667in,height=0.16667in]{images/mastodon.svg}
  \href{https://genomic.social/@miltondp}{@miltondp@genomic.social} \\
  Department of Genetics, Perelman School of Medicine, University of Pennsylvania, Philadelphia, PA 19104, USA
  · Funded by The National Human Genome Research Institute, K99 HG011898; The Eunice Kennedy Shriver National Institute of Child Health and Human Development, R01 HD109765
\item
  \textbf{Casey S. Greene}
  \textsuperscript{\protect\hyperlink{correspondence}{\Letter}}
  \includesvg[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0000-0001-8713-9213}{0000-0001-8713-9213}
  · \includesvg[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/cgreene}{cgreene}
  · \includesvg[width=0.16667in,height=0.16667in]{images/twitter.svg}
  \href{https://twitter.com/GreeneScientist}{GreeneScientist}
  · \includesvg[width=0.16667in,height=0.16667in]{images/mastodon.svg}
  \href{https://genomic.social/@greenescientist}{@greenescientist@genomic.social} \\
  Center for Health AI, University of Colorado School of Medicine, Aurora, CO 80045, USA; Department of Biomedical Informatics, University of Colorado School of Medicine, Aurora, CO 80045, USA
  · Funded by The Gordon and Betty Moore Foundation, GBMF4552; The National Human Genome Research Institute, R01 HG010067; The Eunice Kennedy Shriver National Institute of Child Health and Human Development, R01 HD109765
\end{itemize}

@@

\textbf{Objective:} Investigate how models with advanced natural language processing capabilities can reduce the time-consuming process of writing and revising scholarly manuscripts.\\
\textbf{Materials and Methods:} To this end, we integrate large language models into the Manubot publishing ecosystem to suggest revisions for scholarly text.
Our AI-based revision workflow uses a prompt generator that integrates metadata from the manuscript into prompt templates to generate section-specific instructions for the language model.
Then, the model generates a revised version of each paragraph that the human author can review.
We tested our AI-based revision workflow in three case studies of existing manuscripts, including the present one.\\
\textbf{Results:} Our results suggest that these models can capture the concepts in the scholarly text and produce high-quality revisions that improve clarity.
All changes to the manuscript are tracked using a version control system, providing transparency into the human or machine origin of text.\\
\textbf{Conclusions:} Given the amount of time that researchers put into crafting prose, incorporating large language models into the scholarly writing process can significantly improve the type of knowledge work performed by academics.
It can also help scholars to focus on the most important aspects of their work, such as the novelty of their ideas, and automate the most tedious parts such as adhering to a certain writing style.
Although the use of AI-assisted tools for scientific authoring is controversial, our work focuses on revising text written by humans and provides transparency into the origin of text by tracking changes, which can alleviate concerns about the use of AI in scientific authoring.

@@

Manuscripts have been around for thousands of years, but scientific journals have only been around for about 350 years {[}\protect\hyperlink{ref-F3iZfGUC}{1}{]}.
External peer review, which is used by many journals, is even more recent, having been around for less than 100 years {[}\protect\hyperlink{ref-1HMhNrQq1}{2}{]}.
Most manuscripts are written by humans or teams of humans working together to describe new advances, summarize existing literature, or argue for changes in the status quo.
However, scholarly writing is a time-consuming process where results of a study are presented using a specific style and format.
Academics can sometimes be long-winded in getting to key points, making writing more impenetrable to their audience {[}\protect\hyperlink{ref-19YWsShi0}{3}{]}.

Recent advances in computing capabilities and the widespread availability of text, images, and other data on the internet have laid the foundation for artificial intelligence (AI) models with billions of parameters.
Large language models, in particular, are opening the floodgates to new technologies with the capability to transform how society operates {[}\protect\hyperlink{ref-xq1uEbPa}{4}{]}.
OpenAI's models, for instance, have been trained on vast amounts of data and can generate human-like text {[}\protect\hyperlink{ref-bYOaJHMe}{5}{]}.
These models are based on the transformer architecture which uses self-attention mechanisms to model the complexities of language.
The most well-known of these models is the Generative Pre-trained Transformer 3 (GPT-3), which have been shown to be highly effective for a range of language tasks such as generating text, completing code, and answering questions {[}\protect\hyperlink{ref-bYOaJHMe}{5}{]}.
Scientists are already using these tools to improve scientific writing {[}\protect\hyperlink{ref-Svww2RUh}{6}{]}.
This technology has the potential to revolutionize how scientists write and revise scholarly manuscripts, saving time and effort and enabling researchers to focus on more high-level tasks such as data analysis and interpretation.

We present a novel AI-assisted revision tool that envisions a future where authors collaborate with large language models in the writing of their manuscripts.
This workflow builds on the Manubot infrastructure for scholarly publishing {[}\protect\hyperlink{ref-YuJbg3zO}{7}{]}, a platform designed to enable both individual and large-scale collaborative projects {[}\protect\hyperlink{ref-PZMP42Ak}{8},\protect\hyperlink{ref-10gsAq0o}{9}{]}.
Our workflow involves parsing the manuscript, utilizing a large language model with section-specific prompts for revision, and then generating a set of suggested changes to be integrated into the main document.
These changes are presented to the user through the GitHub interface for review.
To evaluate our workflow, we conducted a case study with three Manubot-authored manuscripts that included sections of varying complexity.
Our findings indicate that, in most cases, the models were able to maintain the original meaning of text, improve the writing style, and even interpret mathematical expressions.
Our AI-assisted writing workflow can be incorporated into any Manubot manuscript, and we anticipate it will help authors more effectively communicate their work.

@@

This study has two objectives:
1) investigate how recently released large language models can reduce the time-consuming process of writing and revising scholarly manuscripts;
2) develop a workflow that integrates these models into the Manubot publishing ecosystem to transparently suggest revisions to improve the quality of the original text.

@@

\hypertarget{overview}{%
\subsubsection{Overview}\label{overview}}

\begin{figure}
\hypertarget{fig:ai_revision}{%
\centering
\includesvg[width=0.75\textwidth,height=\textheight]{images/figure_1.svg}
\caption{\textbf{AI-based revision applied on a Manubot-based manuscript.}
\textbf{a)} A manuscript (written with Manubot) with different sections.
\textbf{b)} The prompt generator integrates metadata using prompt templates to generate section-specific prompts for each paragraph.
If a paragraph belongs to a non-standard section, then a default prompt will be used to perform a basic revision only.
The prompt for the Methods section includes the formatting of equations with identifiers.
All sections' prompts include these instructions: \emph{``the text grammar is correct, spelling errors are fixed, and the text has a clear sentence structure''}, although these are only shown for abstracts.}\label{fig:ai_revision}
}
\end{figure}

We implemented an AI-based revision infrastructure in Manubot {[}\protect\hyperlink{ref-YuJbg3zO}{7}{]}, a tool for collaborative writing of scientific manuscripts.
Manubot integrates with popular version control platforms such as GitHub, allowing authors to easily track changes and collaborate on writing in real time.
Furthermore, Manubot automates the process of generating a formatted manuscript (such as HTML, PDF, DOCX; Figure \ref{fig:ai_revision}a shows the HTML output).
Built on this modern and open paradigm, our AI-based revision software was developed using GitHub Actions, which allows the user to easily trigger an automated revision task on the entire manuscript or specific sections of it.

When the user triggers the action, the manuscript is parsed by section and then by paragraph (Figure \ref{fig:ai_revision}b) and passed to the language model along with a set of custom prompts.
The model then returns a revised version of the text.
Our workflow then uses the GitHub API to generate a new pull request, allowing the user to review and modify the output before merging the changes into the manuscript.
This workflow attributes text to either the human user or to the AI language model, which may be important in light of potential future legal decisions that alter the copyright landscape around the outputs of generative models.

We used the \href{https://openai.com/api/}{OpenAI API} for access to these models.
Since this API incurs a cost with each run that depends on manuscript length, we implemented a workflow in GitHub Actions that can be manually triggered by the user.
Our implementation allows users to tune the costs to their needs by allowing them to select specific sections to be revised instead of the entire manuscript.
Additionally, several model parameters can be adjusted to tune costs even further, such as the language model version (including Davinci and Curie, and potentially newly published ones), how much risk the model will take, or the ``quality'' of the completions.
For instance, using Davinci models (the most complex and capable ones), the cost per run is under \$0.50 for most manuscripts.

@@

Our tools are comprised of Python scripts that perform the AI-based revision (\url{https://github.com/greenelab/manubot-ai-editor}) and a GitHub Actions workflow integrated with Manubot.
To run the workflow, the user must specify the branch that will be revised, select the files/sections of the manuscript (optional), specify the language model to use (\texttt{text-davinci-003} by default), and provide the output branch name.
For more advanced users, it is also possible to change most of the tool's behavior or the language model parameters.

@@

As shown in Figure \ref{fig:ai_revision}b, each API request comprises a prompt (the instructions given to the model) and the paragraph to be revised.
The prompt uses the manuscript title and keywords, so both must be accurate to obtain the best revision outcomes.
The other key component to process a paragraph is its section.
For instance, the abstract is a set of sentences with no citations, whereas a paragraph from the Introduction section has several references to other scientific papers.
A paragraph in the Results section has fewer citations but many references to figures or tables, and must provide enough details about the experiments to understand and interpret the outcomes.
The Methods section is more dependent on the type of paper, but in general it has to provide technical details and sometimes mathematical formulas and equations.
Therefore, we designed section-specific prompts, which we found led to the most useful suggestions.
Figures and tables captions, as well as paragraphs that contain only one or two sentences and less than sixty words, are not processed and are copied directly to the output file.

The section of a paragraph is automatically inferred from the file name using a simple strategy, such as if ``introduction'' or ``methods'' is part of the file name.
If the tool fails to infer a section from the file, then the user is still able to specify which section the file belongs to.
The section can be a standard one (abstract, introduction, results, methods, or discussion) for which a specific prompt is used (Figure \ref{fig:ai_revision}b), or a non-standard one for which a default prompt is used to instruct the model to perform basic revision (minimizing the use of jargon, ensuring text grammar is correct, fixing spelling errors, and making sure the text has a clear sentence structure).

@@

Our AI-based revision workflow uses \href{https://beta.openai.com/docs/guides/completion}{text completion} to process each paragraph.
We tested our tool using Davinci and Curie models, including \texttt{text-davinci-003}, \texttt{text-davinci-edit-001} and \texttt{text-curie-001}.
Davinci models are the most powerful GPT-3 model, whereas Curie ones are less capable but faster and less expensive.
We mainly focused on the completion endpoint, as the edits endpoint is currently in beta.
All models can be fine-tuned using different parameters (see \href{https://beta.openai.com/docs/api-reference/completions}{OpenAI - API Reference}), and the most important ones can be easily adjusted using our tool.

Language models for text completion have a context length that indicates the limit of tokens they can process (tokens are common character sequences in text).
This limit includes the size of the prompt and the paragraph, as well as the maximum number of tokens to generate for the completion (parameter \texttt{max\_tokens}).
For instance, the context length of Davinci models is 4,000 and 2,048 for Curie (see \href{https://beta.openai.com/docs/models/overview}{OpenAI - Models overview}).
Therefore, it is not possible to use the entire manuscript as input, not even entire sections.
To address this limitation, our AI-assisted revision software processes each paragraph of the manuscript with section-specific prompts, as shown in Figure \ref{fig:ai_revision}b.
This approach allows us to process large manuscripts by breaking them into small chunks of text.
However, since the language model only processes a single paragraph from a section, it can potentially lose important context to produce a better output.
Nonetheless, we find that the model still produces high-quality revisions (see \protect\hyperlink{sec:results}{Results}).
Additionally, the maximum number of tokens (parameter \texttt{max\_tokens}) is set as twice the estimated number of tokens in the paragraph (one token approximately represents four characters, see \href{https://beta.openai.com/tokenizer\%5D}{OpenAI - Tokenizer}.
The tool automatically adjusts this parameter and performs the request again if a related error is returned by the API.
The user can also force the tool to either use a fixed value for \texttt{max\_tokens} for all paragraphs, or change the fraction of maximum tokens based on the estimated paragraph size (two by default).

The language models used are stochastic, meaning they generate a different revision for the same input paragraph each time.
This behavior can be adjusted by using the ``sampling temperature'' or ``nucleus sampling'' parameters (we use \texttt{temperature=0.5} by default).
Although we selected default values that worked well across multiple manuscripts, these parameters can be changed to make the model more deterministic.
The user can also instruct the model to generate several completions and select the one with the highest log probability per token, which can improve the quality of the revision.
Our proof-of-concept implementation generates only one completion (parameter \texttt{best\_of=1}) to avoid potentially high costs for the user.
Additionally, our workflow allows the user to process either the entire manuscript or individual sections.
This allows for more cost-effective control while focusing on a single piece of text, wherein the user can run the tool several times and pick the preferred revised text.

@@

We have contributed our workflow (\url{https://github.com/manubot/rootstock/pull/484}) to the standard Manubot template manuscript, which is called rootstock and available at \url{https://github.com/manubot/rootstock}.
Users who wish to use the workflow only need to follow the standard procedures to install Manubot.
The section ``AI-assisted authoring'', in the file \texttt{USAGE.md} of the rootstock repository, explains how to enable the tool.
After that, the workflow (named \texttt{ai-revision}) will be available and ready to use under the Actions tab of the user's manuscript repository.

@@

We evaluated our AI-assisted revision workflow using three GPT-3 models from OpenAI: \texttt{text-davinci-003}, \texttt{text-davinci-edit-001}, and \texttt{text-curie-001}.
The first two are based on the most capable Davinci models (see \href{https://beta.openai.com/docs/models/gpt-3}{OpenAI - GPT-3 models}).
Whereas \texttt{text-davinci-003} is a production-ready model for the completion endpoint, \texttt{text-davinci-edit-001} is used for the edits endpoint and is still in beta.
The latter provides a more natural interface for revising manuscripts, as it takes two inputs: instructions and the text to revise.
Model \texttt{text-curie-001} is faster and cheaper than Davinci models, and is defined as ``very capable'' by its authors (see \href{https://beta.openai.com/docs/models/gpt-3}{OpenAI - GPT-3 models}).

>>>>

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A publishing infrastructure for AI-assisted academic authoring},
  pdfauthor={Milton Pividori, Casey S. Greene},
  pdflang={en-US},
  pdfkeywords={manubot, artificial intelligence, large language models, scholarly publishing, software},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\begin{itemize}
\item
  \textbf{Milton Pividori}
  \includesvg[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0000-0002-3035-4403}{0000-0002-3035-4403}
  · \includesvg[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/miltondp}{miltondp}
  · \includesvg[width=0.16667in,height=0.16667in]{images/twitter.svg}
  \href{https://twitter.com/miltondp}{miltondp}
  · \includesvg[width=0.16667in,height=0.16667in]{images/mastodon.svg}
  \href{https://genomic.social/@miltondp}{@miltondp@genomic.social} \\
  Department of Biomedical Informatics, University of Colorado School of Medicine, Aurora, CO 80045, USA; Department of Genetics, Perelman School of Medicine, University of Pennsylvania, Philadelphia, PA 19104, USA
  · Funded by The National Human Genome Research Institute, K99/R00 HG011898; The Alfred P. Sloan Foundation, G-2023-20989; The Eunice Kennedy Shriver National Institute of Child Health and Human Development, R01 HD109765
\item
  \textbf{Casey S. Greene}
  \textsuperscript{\protect\hyperlink{correspondence}{\Letter}}
  \includesvg[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0000-0001-8713-9213}{0000-0001-8713-9213}
  · \includesvg[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/cgreene}{cgreene}
  · \includesvg[width=0.16667in,height=0.16667in]{images/twitter.svg}
  \href{https://twitter.com/GreeneScientist}{GreeneScientist}
  · \includesvg[width=0.16667in,height=0.16667in]{images/mastodon.svg}
  \href{https://genomic.social/@greenescientist}{@greenescientist@genomic.social} \\
  Center for Health AI, University of Colorado School of Medicine, Aurora, CO 80045, USA; Department of Biomedical Informatics, University of Colorado School of Medicine, Aurora, CO 80045, USA
  · Funded by The Gordon and Betty Moore Foundation, GBMF4552; The National Human Genome Research Institute, R01 HG010067; The Eunice Kennedy Shriver National Institute of Child Health and Human Development, R01 HD109765
\end{itemize}

@@

\textbf{Objective:} In this work, we investigate the use of advanced natural language processing models to reduce the time-consuming process of writing and revising scholarly manuscripts.\\
\textbf{Materials and Methods:} For this purpose, we integrate large language models into the Manubot publishing ecosystem to suggest revisions for scholarly texts.
Our AI-based revision workflow employs a prompt generator that incorporates manuscript metadata into templates, generating section-specific instructions for the language model.
The model then generates revised versions of each paragraph for human authors to review.
We evaluated this methodology through three case studies of existing manuscripts, including the revision of this manuscript.\\
\textbf{Results:} Our results indicate that these models, despite some limitations, can grasp complex academic concepts and enhance text quality.
All changes to the manuscript are tracked using a version control system, ensuring transparency in distinguishing between human- and machine-generated text.\\
\textbf{Conclusions:} Given the significant time researchers invest in crafting prose, incorporating large language models into the scholarly writing process can significantly improve the type of knowledge work performed by academics.
Our approach also enables scholars to concentrate on critical aspects of their work, like the novelty of their ideas, while automating tedious tasks such as adhering to specific writing styles.
Although the use of AI-assisted tools in scientific authoring is controversial, our approach, which focuses on revising human-written text and provides change-tracking transparency, can mitigate concerns regarding AI's role in scientific writing.

@@

The tradition of scholarly manuscripts dates back thousands of years, evolving significantly with the advent of scientific journals approximately 350 years ago {[}\protect\hyperlink{ref-F3iZfGUC}{1}{]}.
External peer review, which is used by many journals, is even more recent, having been around for less than 100 years {[}\protect\hyperlink{ref-1HMhNrQq1}{2}{]}.
Most manuscripts are written by humans or teams of humans working together to describe new advances, summarize existing literature, or argue for changes in the status quo.
However, scholarly writing is a time-consuming process where results of a study are presented using a specific style and format.
Academics can sometimes be long-winded in getting to key points, making writing more impenetrable to their audience {[}\protect\hyperlink{ref-19YWsShi0}{3}{]}.

Recent advances in computing capabilities and the widespread availability of text, images, and other data on the internet have laid the foundation for artificial intelligence (AI) models with billions of parameters.
Large language models (LLMs), in particular, are opening the floodgates to new technologies with the capability to transform how society operates {[}\protect\hyperlink{ref-xq1uEbPa}{4}{]}.
OpenAI's models, for instance, have been trained on vast amounts of data and can generate human-like text {[}\protect\hyperlink{ref-bYOaJHMe}{5}{]}.
These models are based on the transformer architecture which uses self-attention mechanisms to model the complexities of language.
The most well-known of these models is the Generative Pre-trained Transformer (GPT-3 and, more recently, GPT-4), which have been shown to be highly effective for a range of language tasks such as generating text, completing code, and answering questions {[}\protect\hyperlink{ref-bYOaJHMe}{5}{]}.
In the realm of medical informatics, scientists are beginning to explore the utility of these tools in optimizing clinical decision support {[}\protect\hyperlink{ref-gRhoGuC4}{6}{]} or assessing its potential to reduce health disparities {[}\protect\hyperlink{ref-CYB5vhZp}{7}{]}, while also raising concerns about their impact in medical education {[}\protect\hyperlink{ref-h8wInPLE}{8}{]} and the importance of keeping the human aspect central in AI development and application {[}\protect\hyperlink{ref-Z2ek25Ak}{9}{]}.
These tools have been also used in enhancing scientific communication {[}\protect\hyperlink{ref-Svww2RUh}{10}{]}.
This technology has the potential to revolutionize how scientists write and revise scholarly manuscripts, saving time and effort and enabling researchers to focus on more high-level tasks such as data analysis and interpretation.
However, the use of LLMs in research has sparked controversy, primarily due to their propensity to generate plausible yet factually incorrect or misleading information.

In this work, we present a human-centric approach for the use of AI in manuscript writing where scholarly text, initially created by humans, is revised through edit suggestions from LLMs, and then ultimately reviewed and approved by humans.
This approach mitigates the risk of generating misleading information while still providing the benefits of AI-assisted writing.
We developed an AI-assisted revision tool that implements this approach and builds on the Manubot infrastructure for scholarly publishing {[}\protect\hyperlink{ref-YuJbg3zO}{11}{]}, a platform designed to enable both individual and large-scale collaborative projects {[}\protect\hyperlink{ref-PZMP42Ak}{12},\protect\hyperlink{ref-10gsAq0o}{13}{]}.
Our tool, named the Manubot AI Editor, parses the manuscript, utilizes an LLM with section-specific prompts for revision, and then generates a set of suggested changes to be integrated into the main document.
These changes are presented to the user through the GitHub interface for review.
During prompt engineering, we developed unit tests to ensure that a minimum set of quality measures are met by the AI revisions.
For end-to-end evaluation, we manually reviewed the AI revisions on three Manubot-authored manuscripts that included sections of varying complexity.
Our findings indicate that, in most cases, the models were able to maintain the original meaning of text, improve the writing style, and even interpret mathematical expressions.
Officially part of the Manubot platform, our Manubot AI Editor can be readily incorporated into Manubot-based manuscripts, and we anticipate it will help authors more effectively communicate their work.

@@

This study has two objectives:
1) investigate how recently released large language models can reduce the time-consuming process of writing and revising scholarly manuscripts;
2) develop an approach that integrates these models into the Manubot publishing ecosystem to transparently suggest revisions to improve the quality of the original text.

@@

We propose a human-centric approach for the use of AI in manuscript writing, which consists of the following steps:
1) human authors write the manuscript content;
2) an LLM revises the manuscript, generating a set of suggested changes;
3) human authors review the suggested changes, and the approved edits are then integrated into the manuscript.
By focusing on human review, this approach attempts to mitigate the risk of generating incorrect or misleading information.
To implement this human-centric approach, we developed a tool called the Manubot AI Editor, which is part of the Manubot infrastructure for scholarly publishing {[}\protect\hyperlink{ref-YuJbg3zO}{11}{]}.

\hypertarget{overview-of-the-manubot-ai-editor}{%
\subsubsection{Overview of the Manubot AI Editor}\label{overview-of-the-manubot-ai-editor}}

\begin{figure}
\hypertarget{fig:ai_revision}{%
\centering
\includesvg[width=0.75\textwidth,height=\textheight]{images/figure_1.svg}
\caption{\textbf{AI-based revision applied on a Manubot-based manuscript.}
\textbf{a)} A manuscript (written with Manubot) with different sections.
\textbf{b)} The prompt generator integrates metadata using prompt templates to generate section-specific prompts for each paragraph.
If a paragraph belongs to a non-standard section, then a default prompt will be used to perform a basic revision only.
The prompt for the Methods section includes the formatting of equations with identifiers.
All sections' prompts include these instructions: \emph{``the text grammar is correct, spelling errors are fixed, and the text has a clear sentence structure''}, although these are only shown for abstracts.
Our tool allows the user to provide a custom prompt instead of using the default ones shown here.}\label{fig:ai_revision}
}
\end{figure}

The Manubot AI Editor is an AI-based revision infrastructure built into Manubot {[}\protect\hyperlink{ref-YuJbg3zO}{11}{]}, a tool for collaborative writing of scientific manuscripts.
Manubot integrates with popular version control platforms such as GitHub, allowing authors to easily track changes and collaborate on writing in real time.
Furthermore, Manubot automates the process of generating a formatted manuscript (such as HTML, PDF, DOCX; Figure \ref{fig:ai_revision}a shows the HTML output).
Built on this modern and open paradigm, our Manubot AI Editor (\url{https://github.com/manubot/manubot-ai-editor}) includes three components:
1) a Python library that provides classes and functions to read the manuscript content and its metadata, calls the LLM for automatic text revision, and writes the results back;
2) a GitHub Actions workflow that uses our Python library within GitHub to preserve provenance information for transparency;
3) a prompt generator that integrates the manuscript's metadata using prompt templates to generate section-specific prompts for each paragraph (Figure \ref{fig:ai_revision}b).

The GitHub Actions workflow allows the user to easily trigger an automated revision task on the entire manuscript or specific sections of it.
When the user triggers the action, the manuscript is parsed by section and then by paragraph (Figure \ref{fig:ai_revision}b) that are passed to the language model along with a set of custom prompts.
The model then returns a revised version of the text.
Our workflow then uses the GitHub API to generate a new pull request, allowing the user to review and modify the output before merging the changes into the manuscript.
This workflow attributes text to either the human user or to the AI language model, which may be important in light of potential future legal decisions that alter the copyright landscape around the outputs of generative models.

We used the \href{https://openai.com/api/}{OpenAI API} for access to these models.
Since this API incurs a cost with each run that depends on manuscript length, we implemented a workflow in GitHub Actions that can be manually triggered by the user.
Our implementation allows users to tune the costs to their needs by allowing them to select specific sections to be revised instead of the entire manuscript.
Additionally, several model parameters can be adjusted to tune costs even further, such as the language model version (including Davinci and Curie, the current GPT-3.5 Turbo and GPT-4, and potentially newly published ones), how much risk the model will take, or the ``quality'' of the completions.
For instance, using Davinci models, the cost per run is under \$0.50 for most manuscripts.

@@

To run the workflow, the user must specify the branch that will be revised, select the files/sections of the manuscript (optional), specify the language model to use (\texttt{text-davinci-003} by default), an optional custom prompt (section-specific prompts are used by default), and provide the output branch name.
For more advanced users, it is also possible to change most of the tool's behavior or the language model parameters.

@@

As shown in Figure \ref{fig:ai_revision}b, each API request comprises a prompt (the instructions given to the model) and the paragraph to be revised.
Unless the user specifies a custom prompt, the tool will use a section-specific prompt generator that incorporates the manuscript title and keywords, so both must be accurate to obtain the best revision outcomes.
The other key component to process a paragraph is its section.
For instance, the abstract is a set of sentences with no citations, whereas a paragraph from the Introduction section has several references to other scientific papers.
A paragraph in the Results section has fewer citations but many references to figures or tables, and must provide enough details about the experiments to understand and interpret the outcomes.
The Methods section is more dependent on the type of paper, but in general it has to provide technical details and sometimes mathematical formulas and equations.
Therefore, we designed section-specific prompts, which we found led to the most useful suggestions.
Figures and tables captions, as well as paragraphs that contain only one or two sentences and less than sixty words, are not processed and are copied directly to the output file.

The section of a paragraph is automatically inferred from the file name using a simple strategy, such as if ``introduction'' or ``methods'' is part of the file name.
If the tool fails to infer a section from the file, then the user is still able to specify which section the file belongs to.
The section can be a standard one (abstract, introduction, results, methods, or discussion) for which a specific prompt is used (Figure \ref{fig:ai_revision}b), or a non-standard one for which a default prompt is used to instruct the model to perform basic revision (\emph{``minimizing the use of jargon, ensuring text grammar is correct, fixing spelling errors, and making sure the text has a clear sentence structure''}).

@@

Our AI-based revision workflow uses the \href{https://platform.openai.com/docs/guides/text-generation/chat-completions-api}{Chat Completions API} to process each paragraph.
We tested our tool using Davinci and Curie models, including \texttt{text-davinci-003}, \texttt{text-davinci-edit-001} and \texttt{text-curie-001}.
Within the GPT-3 family, Davinci models are the most powerful, whereas Curie ones are less capable but faster and less expensive.
All models can be fine-tuned using different parameters (see \href{https://platform.openai.com/docs/api-reference/chat/create}{OpenAI - API Reference}), and the most important ones can be easily adjusted using our tool.

Language models for text completion have a context length that indicates the limit of tokens they can process (tokens are common character sequences in text).
This limit includes the size of the prompt and the paragraph, as well as the maximum number of tokens to generate for the completion (parameter \texttt{max\_tokens}).
For instance, the context length of Davinci models is 4,000 and 2,048 for Curie (see \href{https://platform.openai.com/docs/models/gpt-3}{OpenAI - Models overview}).
To make sure we never surpass this context length, our AI-assisted revision software processes each paragraph of the manuscript with section-specific prompts, as shown in Figure \ref{fig:ai_revision}b.
This approach allows us to process large manuscripts by breaking them into small chunks of text.
However, since the language model only processes a single paragraph from a section, it can potentially lose important context to produce a better output.
Nonetheless, we find that the model still produces high-quality revisions (see \protect\hyperlink{sec:results}{Results}).
Additionally, the maximum number of tokens (parameter \texttt{max\_tokens}) is set as twice the estimated number of tokens in the paragraph (one token approximately represents four characters, see \href{https://platform.openai.com/tokenizer\%5D}{OpenAI - Tokenizer}.
The tool automatically adjusts this parameter and performs the request again if a related error is returned by the API.
The user can also force the tool to either use a fixed value for \texttt{max\_tokens} for all paragraphs, or change the fraction of maximum tokens based on the estimated paragraph size (two by default).

The language models used are stochastic, meaning they generate a different revision for the same input paragraph each time.
This behavior can be adjusted by using the ``sampling temperature'' or ``nucleus sampling'' parameters (we use \texttt{temperature=0.5} by default).
Although we selected default values that worked well across multiple manuscripts, these parameters can be changed to make the model more deterministic.
The user can also instruct the model to generate several completions and select the one with the highest log probability per token, which can improve the quality of the revision.
Our implementation generates only one completion (parameter \texttt{best\_of=1}) to avoid potentially high costs for the user.
Additionally, our workflow allows the user to process either the entire manuscript or individual sections.
This allows for more cost-effective control while focusing on a single piece of text, wherein the user can run the tool several times and pick the preferred revised text.

@@

The Manubot AI Editor is part of the standard Manubot template manuscript, which is called rootstock and available at \url{https://github.com/manubot/rootstock}.
Users who wish to use the workflow only need to follow the standard procedures to install Manubot.
The section ``AI-assisted authoring'', in the file \texttt{USAGE.md} of the rootstock repository, explains how to enable the tool.
After that, the workflow (named \texttt{ai-revision}) will be available and ready to use under the Actions tab of the user's manuscript repository.

@@

Assessing the performance of text generation tasks is challenging, and this is specially true for automatic revisions of scientific content.
In this context, we need to make sure the revision does not change the original meaning or does not introduce incorrect or misleading information.
For this reason, our approach emphasizes human assessments of the revisions to mitigate these issues, and we followed the same procedure in evaluating our tool.
We used three manuscripts of our own authorship (see below), which allowed us to more objetively assess changes in the original meaning and whether revisions kept important details.
During the prompt engineering phase (see below), we also used a unit testing framework to ensure that the revisions produced by our prompts met a minimum set of quality measures.

\hypertarget{language-models}{%
\paragraph{Language models}\label{language-models}}

We evaluated our AI-assisted revision workflow using three GPT-3 models from OpenAI: \texttt{text-davinci-003}, \texttt{text-davinci-edit-001}, and \texttt{text-curie-001}.
The first two are based on the most capable GPT-3 Davinci models (see \href{https://platform.openai.com/docs/models/gpt-3}{OpenAI - GPT-3 models}).
Whereas \texttt{text-davinci-003} is a production-ready model for the completion endpoint, \texttt{text-davinci-edit-001} is used for the edits endpoint and was still in beta at the time of testing.
The latter provides a more natural interface for revising manuscripts, as it takes two inputs: instructions and the text to revise.
Model \texttt{text-curie-001} is faster and cheaper than Davinci models, and is defined as ``very capable'' by its authors (see \href{https://platform.openai.com/docs/models/gpt-3}{OpenAI - GPT-3 models}).

\hypertarget{manuscripts}{%
\paragraph{Manuscripts}\label{manuscripts}}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1467}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2822}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3767}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1944}}@{}}
\caption{\textbf{Manuscripts used to evaluate the AI-based revision workflow.} The title and keywords of a manuscript are used in prompts for revising paragraphs. IDs are used in the text to refer to them. \label{tbl:manuscripts}}\label{tbl:manuscripts}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Manuscript ID
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
GitHub URL
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Title
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Keywords
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Manuscript ID
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
GitHub URL
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Title
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Keywords
\end{minipage} \\
\midrule()
\endhead
CCC & \href{https://github.com/greenelab/ccc-manuscript}{greenelab/ccc-manuscript} & An efficient not-only-linear correlation coefficient based on machine learning & correlation coefficient, nonlinear relationships, gene expression \\
PhenoPLIER & \href{https://github.com/greenelab/phenoplier_manuscript}{greenelab/phenoplier\_manuscript} & Projecting genetic associations through gene expression patterns highlights disease etiology and drug mechanisms & genetic studies, functional genomics, gene co-expression, therapeutic targets, drug repurposing, clustering of complex traits \\
Manubot-AI & \href{https://github.com/greenelab/manubot-gpt-manuscript}{greenelab/manubot-gpt-manuscript} & A publishing infrastructure for AI-assisted academic authoring & manubot, artificial intelligence, scholarly publishing, software \\
\bottomrule()
\end{longtable}
